\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-01 coq-example.tex
Add diagram for abs-int-example.

\subsection{Abstract Interpreters for \texttt{Imp}}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for both concrete interpreters and
abstract interpreters.
Thanks to the compilation to Coq, we can directly run the resulting
abstract interpreters.

We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful. They are here to show family inheritance is useful, and can extend stuff to all sorts of things. I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family? What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost. This is just simulating people starting with bare (big-step) operational semantic. Because they may not working on abstract interpretation from the first place. Those people may extend "Lang" with other stuff like another type-safety proof or something.}

The first family "Lang" define the big step operator (using fuel) of a
while language, with expression. However, it postulates the definition of "State" (environment mapping) used by Lang. It takes about 200+ LOCs. 
% We
% can
\YZ{In general, in a result section like this, you want to show what
you *have done* rather than what you *can do*.}\EDJreply{Ok. deleted. I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
% swap the implementation of "State" to have different memory/state
% management by extending family "Lang".
\YZ{What is 'different
memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary. I instantiate it with function(closure) as environment mapping in "LangRun". It is possible to use other mapping like Coq's primitive dictionary.}

The second family "LangwAbs" define the abstract interpreter for it,
together with a bunch of postulates, to prove the correctness of the
abstract interpreter. It takes about 300+ LOCs. This example shows how
to use family inheritance to extend a concrete interpreter with abstract
interpreters. This example also shows in this family polymoprhism
framework we can reason about "Lang" with computation details being
abstracted.

The third Family "LangMore" extend LangwAbs with nat constant and
addition, and if-then-else control flow, of course retaining all the
soundness theorem. It takes about 200+ LOCs. This is one example of
using family inheritance to support new language feature, and compatible
with the existent reasoning.

The fourth family "LangRun" instantiate some of the postulates of the
LangMore and recover computational information of Lang. The concrete
interpreter is doing as expected; the abstract interpreter is expected
to act as a type-analyzer. It takes about 200+ LOCs. This is one example
of instantiate the detail computation. Thanks to the posulation we have done in "Lang", now we instantiate that postulation with function type "Ident -> option exp" as the "State" (environment mapping) for simplicity.

And at the end, we compile LangRun into module and use Coq to run it on
some simple queries. This "LangRun" also illustrates that we still have
computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}

>>>>

\subsection{Abstract Interpreters for \texttt{Imp}}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for both concrete interpreters and
abstract interpreters.
Thanks to the compilation to Coq, we can directly run the resulting
abstract interpreters.

We construct four families, in one inheritance chain, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful. They are here to show family inheritance is useful, and can extend stuff to all sorts of things. I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family? What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost. This is just simulating people starting with bare (big-step) operational semantic. Because they may not working on abstract interpretation from the first place. Those people may extend "Lang" with other stuff like another type-safety proof or something.}

The first family "Lang" define the big step operator (using fuel) of a
while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200+ LOCs. 
% We
% can
\YZ{In general, in a result section like this, you want to show what
you *have done* rather than what you *can do*.}\EDJreply{Ok. deleted. I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
% swap the implementation of "State" to have different memory/state
% management by extending family "Lang".
\YZ{What is 'different
memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary. I instantiate it with function(closure) as environment mapping in "LangRun". It is possible to use other mapping like Coq's primitive dictionary.}

The second family "LangwAbs" define the abstract interpreter for "Lang" in "LangwAbs.AbsInterpreter",
together with a bunch of postulates as a big sealed family "absStateData". Then we prove the correctness of the
abstract interpretation function "analyze". It takes about 300+ LOCs. This example shows how
to use family inheritance to extend a concrete interpreter with abstract
interpreters. 
% This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.

The third Family "LangMore" extend "LangwAbs" with nat constant and
addition, and if-then-else control flow, of course retaining all the
soundness theorem. It takes about 200+ LOCs. This is one example of
using family inheritance to support new language feature, and compatible
with the existent reasoning in "LangwAbs.AbsInterpreter".

The fourth family "LangRun" instantiate some of the postulates of the
"LangMore" and recover computational information of "analyze". The "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
to act as a type-analyzer. It takes about 200+ LOCs. This is one example
of instantiate the detail computation of abstract interpretation.

And at the end, we compile LangRun into module and use Coq to run it on
some simple queries. This "LangRun" also illustrates that we still have
computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done. Please check}

\begin{figure}
  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
  \caption{Inheritance Diagram for Abstract Interpretation Example. Black arrows are inheritance. Purple arrow are extending. Green arrows are overriding. Grey box are sealed families.}\label{fig:abstract-interpretation-example}
\end{figure}

### 2022-07-01 coqimpl.tex
Add diagram for abs-int-example.

>>>>

\begin{figure}[!htb]
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC.
FInductive ty : Set :=
  | unit : ty
  | arrow : ty -> ty -> ty.
FInductive term : Set := 
  | var : id -> term 
  | lam : id -> term -> term ...
 Family subst_internal.
 EndFamily.

  : id -> term -> term := ...
Inductive has_type 
  : context -> term -> type := ...
Proposition subst_lemma :
  has_type (G; x ↦ A) t T ->
  has_type G u A ->
  has_type G (subst x u t) T.
Inductive step : term -> term -> Prop 
  := ...
(* ... and more, end with type safety *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{0.47\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC_bool extends STLC.
Extend FInductive ty : Set :=
  | bool : ty.

Extend FInductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term
Extend Fixpoint subst (* Need to handle new term *)
  : id -> term -> term := ...
Extend Inductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
Extend Proposition subst_lemma :
  ... (* Need to prove extra cases *)

Extend Inductive step : term -> term -> Prop 
  := ... (* Need to expand this binary relation *)
(* ... and more extension *)
EndFamily.
\end{lstlisting}
  \end{minipage}
  \caption{Example STLC and its extension, in designed syntax}\label{fig:STLC-example2}
\end{figure}

### 2022-07-01 syntax.tex
Add diagram for abs-int-example.

Inheritance are
judgement so naturally second class citizen data (i.e. not something
function can return and cannot be assigned to variables), and it looks
very much like a function that transform a given linkage. Like plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and an overriding term can only refer to fields before it.

>>>>

Inheritance are
judgement so naturally second class citizen data (i.e. not something
function can return and cannot be assigned to variables), and it looks
very much like a function that transform a given linkage. Similar to plugin implementation and \ruleref{Lnkg Add}, we resolve \ref{chg:consistency} by making sure corresponding fields are in corresponding context during inheritance, and thus an overriding term can only refer to fields before the overriden field.
